\newpage
\begin{thebibliography}{1}
\bibitem[\protect\citeauthoryear{Agrawal, Gans, and Goldfarb}{Agrawal et~al.}{2019}]{agrawal2019economics}
\textsc{Agrawal, A., J.~Gans, and A.~Goldfarb} (2019): \emph{The economics of artificial intelligence: an agenda}, University of Chicago Press.

\bibitem[\protect\citeauthoryear{Amodei, Olah, Steinhardt, Christiano, Schulman, and Man{\'e}}{Amodei et~al.}{2016}]{amodei2016concrete}
\textsc{Amodei, D., C.~Olah, J.~Steinhardt, P.~Christiano, J.~Schulman, and D.~Man{\'e}} (2016): \enquote{Concrete problems in AI safety,} \emph{arXiv preprint arXiv:1606.06565}.

\bibitem[\protect\citeauthoryear{Argote and Miron-Spektor}{Argote and Miron-Spektor}{2011}]{argote2011organizational}
\textsc{Argote, L. and E.~Miron-Spektor} (2011): \enquote{Organizational learning: From experience to knowledge,} \emph{Organization science}, 22, 1123--1137.

\bibitem[\protect\citeauthoryear{Arrow}{Arrow}{1962}]{arrow1962economic}
\textsc{Arrow, K.~J.} (1962): \enquote{The economic implications of learning by doing,} \emph{The review of economic studies}, 29, 155--173.

\bibitem[\protect\citeauthoryear{Arrow, Chenery, Minhas, and Solow}{Arrow et~al.}{1961}]{arrow1961capital}
\textsc{Arrow, K.~J., H.~B. Chenery, B.~S. Minhas, and R.~M. Solow} (1961): \enquote{Capital-labor substitution and economic efficiency,} \emph{The review of Economics and Statistics}, 225--250.

\bibitem[\protect\citeauthoryear{Baddeley}{Baddeley}{1992}]{baddeley1992working}
\textsc{Baddeley, A.} (1992): \enquote{Working memory,} \emph{Science}, 255, 556--559.

\bibitem[\protect\citeauthoryear{Baumeister, Bratslavsky, Muraven, and Tice}{Baumeister et~al.}{2018}]{baumeister2018ego}
\textsc{Baumeister, R.~F., E.~Bratslavsky, M.~Muraven, and D.~M. Tice} (2018): \enquote{Ego depletion: Is the active self a limited resource?} in \emph{Self-regulation and self-control}, Routledge, 16--44.

\bibitem[\protect\citeauthoryear{Becker}{Becker}{1965}]{becker1965theory}
\textsc{Becker, G.~S.} (1965): \enquote{A Theory of the Allocation of Time,} \emph{The economic journal}, 75, 493--517.

\bibitem[\protect\citeauthoryear{Bellman}{Bellman}{1952}]{bellman1952theory}
\textsc{Bellman, R.} (1952): \enquote{On the theory of dynamic programming,} \emph{Proceedings of the national Academy of Sciences}, 38, 716--719.

\bibitem[\protect\citeauthoryear{Borghini, Vecchiato, Toppi, Astolfi, Maglione, Isabella, Caltagirone, Kong, Wei, Zhou et~al.}{Borghini et~al.}{2012}]{borghini2012assessment}
\textsc{Borghini, G., G.~Vecchiato, J.~Toppi, L.~Astolfi, A.~Maglione, R.~Isabella, C.~Caltagirone, W.~Kong, D.~Wei, Z.~Zhou, et~al.} (2012): \enquote{Assessment of mental fatigue during car driving by using high resolution EEG activity and neurophysiologic indices,} in \emph{2012 annual international conference of the IEEE engineering in medicine and biology society}, IEEE, 6442--6445.

\bibitem[\protect\citeauthoryear{Buser and Peter}{Buser and Peter}{2012}]{buser2012multitasking}
\textsc{Buser, T. and N.~Peter} (2012): \enquote{Multitasking,} \emph{Experimental economics}, 15, 641--655.

\bibitem[\protect\citeauthoryear{Bzdok, Krzywinski, and Altman}{Bzdok et~al.}{2018}]{bzdok2018machine}
\textsc{Bzdok, D., M.~Krzywinski, and N.~Altman} (2018): \enquote{Machine learning: supervised methods,} \emph{Nature methods}, 15, 5.

\bibitem[\protect\citeauthoryear{Caplin}{Caplin}{2016}]{caplin2016measuring}
\textsc{Caplin, A.} (2016): \enquote{Measuring and modeling attention,} \emph{Annual Review of Economics}, 8, 379--403.

\bibitem[\protect\citeauthoryear{Caplin and Dean}{Caplin and Dean}{2015}]{caplin2015revealed}
\textsc{Caplin, A. and M.~Dean} (2015): \enquote{Revealed preference, rational inattention, and costly information acquisition,} \emph{American Economic Review}, 105, 2183--2203.

\bibitem[\protect\citeauthoryear{Cappelli and Rogovsky}{Cappelli and Rogovsky}{2023}]{cappelli2023artificial}
\textsc{Cappelli, P. and N.~G. Rogovsky} (2023): \emph{Artificial intelligence in human resource management: A challenge for the human-centred agenda?}, 95, ILO Working Paper.

\bibitem[\protect\citeauthoryear{Cobb and Douglas}{Cobb and Douglas}{1928}]{cobb1928theory}
\textsc{Cobb, C.~W. and P.~H. Douglas} (1928): \enquote{A theory of production,} .

\bibitem[\protect\citeauthoryear{Dauphin, Pascanu, Gulcehre, Cho, Ganguli, and Bengio}{Dauphin et~al.}{2014}]{dauphin2014identifying}
\textsc{Dauphin, Y.~N., R.~Pascanu, C.~Gulcehre, K.~Cho, S.~Ganguli, and Y.~Bengio} (2014): \enquote{Identifying and attacking the saddle point problem in high-dimensional non-convex optimization,} \emph{Advances in neural information processing systems}, 27.

\bibitem[\protect\citeauthoryear{Davenport, Ronanki et~al.}{Davenport et~al.}{2018}]{davenport2018artificial}
\textsc{Davenport, T.~H., R.~Ronanki, et~al.} (2018): \enquote{Artificial intelligence for the real world,} \emph{Harvard business review}, 96, 108--116.

\bibitem[\protect\citeauthoryear{Deck and Jahedi}{Deck and Jahedi}{2015}]{deck2015effect}
\textsc{Deck, C. and S.~Jahedi} (2015): \enquote{The effect of cognitive load on economic decision making: A survey and new experiments,} \emph{European Economic Review}, 78, 97--119.

\bibitem[\protect\citeauthoryear{Dehm, Nguyen, and Stadje}{Dehm et~al.}{2023}]{dehm2023non}
\textsc{Dehm, C., T.~Nguyen, and M.~Stadje} (2023): \enquote{Non-concave expected utility optimization with uncertain time horizon,} \emph{Applied Mathematics \& Optimization}, 88, 65.

\bibitem[\protect\citeauthoryear{Dewan and Neligh}{Dewan and Neligh}{2020}]{dewan2020estimating}
\textsc{Dewan, A. and N.~Neligh} (2020): \enquote{Estimating information cost functions in models of rational inattention,} \emph{Journal of Economic Theory}, 187, 105011.

\bibitem[\protect\citeauthoryear{Dixit}{Dixit}{1990}]{dixit1990optimization}
\textsc{Dixit, A.~K.} (1990): \emph{Optimization in economic theory}, OUP UK.

\bibitem[\protect\citeauthoryear{Drichoutis and Nayga~Jr}{Drichoutis and Nayga~Jr}{2020}]{drichoutis2020economic}
\textsc{Drichoutis, A.~C. and R.~M. Nayga~Jr} (2020): \enquote{Economic rationality under cognitive load,} \emph{The Economic Journal}, 130, 2382--2409.

\bibitem[\protect\citeauthoryear{Elsken, Metzen, and Hutter}{Elsken et~al.}{2019}]{elsken2019neural}
\textsc{Elsken, T., J.~H. Metzen, and F.~Hutter} (2019): \enquote{Neural architecture search: A survey,} \emph{Journal of Machine Learning Research}, 20, 1--21.

\bibitem[\protect\citeauthoryear{Engle}{Engle}{2018}]{engle2018working}
\textsc{Engle, R.~W.} (2018): \enquote{Working memory and executive attention: A revisit,} \emph{Perspectives on psychological science}, 13, 190--193.

\bibitem[\protect\citeauthoryear{Fl{\aa}m, Jongen, and Stein}{Fl{\aa}m et~al.}{2008}]{flaam2008slopes}
\textsc{Fl{\aa}m, S.~D., H.~T. Jongen, and O.~Stein} (2008): \enquote{Slopes of shadow prices and Lagrange multipliers,} \emph{Optimization Letters}, 2, 143--155.

\bibitem[\protect\citeauthoryear{Frey and Stutzer}{Frey and Stutzer}{2005}]{frey2005beyond}
\textsc{Frey, B.~S. and A.~Stutzer} (2005): \enquote{Beyond outcomes: measuring procedural utility,} \emph{Oxford Economic Papers}, 57, 90--111.

\bibitem[\protect\citeauthoryear{Furman and Seamans}{Furman and Seamans}{2019}]{furman2019ai}
\textsc{Furman, J. and R.~Seamans} (2019): \enquote{AI and the Economy,} \emph{Innovation policy and the economy}, 19, 161--191.

\bibitem[\protect\citeauthoryear{Giray}{Giray}{2023}]{giray2023prompt}
\textsc{Giray, L.} (2023): \enquote{Prompt engineering with ChatGPT: a guide for academic writers,} \emph{Annals of biomedical engineering}, 51, 2629--2633.

\bibitem[\protect\citeauthoryear{Grandl, Ananthanarayanan, Kandula, Rao, and Akella}{Grandl et~al.}{2014}]{grandl2014multi}
\textsc{Grandl, R., G.~Ananthanarayanan, S.~Kandula, S.~Rao, and A.~Akella} (2014): \enquote{Multi-resource packing for cluster schedulers,} \emph{ACM SIGCOMM Computer Communication Review}, 44, 455--466.

\bibitem[\protect\citeauthoryear{Gu, Alazab, Lin, Zhang, and Huang}{Gu et~al.}{2022}]{gu2022ai}
\textsc{Gu, B., M.~Alazab, Z.~Lin, X.~Zhang, and J.~Huang} (2022): \enquote{AI-enabled task offloading for improving quality of computational experience in ultra dense networks,} \emph{ACM Transactions on Internet Technology (TOIT)}, 22, 1--17.

\bibitem[\protect\citeauthoryear{Hagger, Wood, Stiff, and Chatzisarantis}{Hagger et~al.}{2010}]{hagger2010ego}
\textsc{Hagger, M.~S., C.~Wood, C.~Stiff, and N.~L. Chatzisarantis} (2010): \enquote{Ego depletion and the strength model of self-control: a meta-analysis.} \emph{Psychological bulletin}, 136, 495.

\bibitem[\protect\citeauthoryear{Hallak and Teboulle}{Hallak and Teboulle}{2020}]{hallak2020finding}
\textsc{Hallak, N. and M.~Teboulle} (2020): \enquote{Finding second-order stationary points in constrained minimization: A feasible direction approach,} \emph{Journal of Optimization Theory and Applications}, 186, 480--503.

\bibitem[\protect\citeauthoryear{He, Zhang, Ren, and Sun}{He et~al.}{2016}]{he2016deep}
\textsc{He, K., X.~Zhang, S.~Ren, and J.~Sun} (2016): \enquote{Deep residual learning for image recognition,} in \emph{Proceedings of the IEEE conference on computer vision and pattern recognition}, 770--778.

\bibitem[\protect\citeauthoryear{Heath}{Heath}{2019}]{heath2019prediction}
\textsc{Heath, D.~R.} (2019): \enquote{Prediction machines: the simple economics of artificial intelligence: by Ajay Agrawal, Joshua Gans and Avi Goldfarb, Published in 2018 by Harvard Business Review Press, 272 pp., 30.00(hardcover),KindleEdition: 16.19, ISBN: 978-1-633695672,} .

\bibitem[\protect\citeauthoryear{Hoffmann, Boysel, Nagle, Peng, and Xu}{Hoffmann et~al.}{2024}]{hoffmann2024generative}
\textsc{Hoffmann, M., S.~Boysel, F.~Nagle, S.~Peng, and K.~Xu} (2024): \enquote{Generative AI and the Nature of Work,} \emph{Harvard Business School Strategy Unit Working Paper}, 25--021.

\bibitem[\protect\citeauthoryear{Holmstrom and Milgrom}{Holmstrom and Milgrom}{1991}]{holmstrom1991multitask}
\textsc{Holmstrom, B. and P.~Milgrom} (1991): \enquote{Multitask principal--agent analyses: Incentive contracts, asset ownership, and job design,} \emph{The Journal of Law, Economics, and Organization}, 7, 24--52.

\bibitem[\protect\citeauthoryear{Inzlicht, Werner, Briskin, and Roberts}{Inzlicht et~al.}{2021}]{inzlicht2021integrating}
\textsc{Inzlicht, M., K.~M. Werner, J.~L. Briskin, and B.~W. Roberts} (2021): \enquote{Integrating models of self-regulation,} \emph{Annual review of psychology}, 72, 319--345.

\bibitem[\protect\citeauthoryear{Jie and Yan}{Jie and Yan}{2021}]{jie2021computing}
\textsc{Jie, T. and G.~Yan} (2021): \enquote{Computing shadow prices with multiple Lagrange multipliers.} \emph{Journal of Industrial \& Management Optimization}, 17.

\bibitem[\protect\citeauthoryear{Kahneman}{Kahneman}{1973}]{kahneman1973attention}
\textsc{Kahneman, D.} (1973): \enquote{Attention and effort,} .

\bibitem[\protect\citeauthoryear{Kaplan, McCandlish, Henighan, Brown, Chess, Child, Gray, Radford, Wu, and Amodei}{Kaplan et~al.}{2020}]{kaplan2020scaling}
\textsc{Kaplan, J., S.~McCandlish, T.~Henighan, T.~B. Brown, B.~Chess, R.~Child, S.~Gray, A.~Radford, J.~Wu, and D.~Amodei} (2020): \enquote{Scaling laws for neural language models,} \emph{arXiv preprint arXiv:2001.08361}.

\bibitem[\protect\citeauthoryear{Kim and Zauberman}{Kim and Zauberman}{2019}]{kim2019effect}
\textsc{Kim, K. and G.~Zauberman} (2019): \enquote{The effect of music tempo on consumer impatience in intertemporal decisions,} \emph{European Journal of Marketing}, 53, 504--523.

\bibitem[\protect\citeauthoryear{L{\"a}mmermann, Hofmann, and Urbach}{L{\"a}mmermann et~al.}{2024}]{lammermann2024managing}
\textsc{L{\"a}mmermann, L., P.~Hofmann, and N.~Urbach} (2024): \enquote{Managing artificial intelligence applications in healthcare: Promoting information processing among stakeholders,} \emph{International Journal of Information Management}, 75, 102728.

\bibitem[\protect\citeauthoryear{Li, Wang, Lu, and Wang}{Li et~al.}{2022}]{li2022optimizing}
\textsc{Li, M., T.~Wang, W.~Lu, and M.~Wang} (2022): \enquote{Optimizing the systematic characteristics of online learning systems to enhance the continuance intention of Chinese college students,} \emph{Sustainability}, 14, 11774.

\bibitem[\protect\citeauthoryear{Loewenstein}{Loewenstein}{1992}]{loewenstein1992anomalies}
\textsc{Loewenstein, G.} (1992): \enquote{Anomalies of Intertemporal Choice: Evidence and Interpretation,} \emph{Russell Sage}.

\bibitem[\protect\citeauthoryear{Loewenstein and Wojtowicz}{Loewenstein and Wojtowicz}{2023}]{loewenstein2023economics}
\textsc{Loewenstein, G. and Z.~Wojtowicz} (2023): \enquote{The Economics of Attention,} \emph{Available at SSRN 4368304}.

\bibitem[\protect\citeauthoryear{Ma{\'c}kowiak, Mat{\v{e}}jka, and Wiederholt}{Ma{\'c}kowiak et~al.}{2023}]{mackowiak2023rational}
\textsc{Ma{\'c}kowiak, B., F.~Mat{\v{e}}jka, and M.~Wiederholt} (2023): \enquote{Rational inattention: A review,} \emph{Journal of Economic Literature}, 61, 226--273.

\bibitem[\protect\citeauthoryear{Mallen and Belrose}{Mallen and Belrose}{2024}]{mallen2024balancing}
\textsc{Mallen, A. and N.~Belrose} (2024): \enquote{Balancing Label Quantity and Quality for Scalable Elicitation,} \emph{arXiv preprint arXiv:2410.13215}.

\bibitem[\protect\citeauthoryear{Mas-Colell, Whinston, and Green}{Mas-Colell et~al.}{1995}]{RePEc:oxp:obooks:9780195102680}
\textsc{Mas-Colell, A., M.~D. Whinston, and J.~R. Green} (1995): \emph{{Microeconomic Theory}}, no. 9780195102680 in OUP Catalogue, Oxford University Press.

\bibitem[\protect\citeauthoryear{Mirhosseini, Sadrosadati, Soltani, Sarbazi-Azad, and Wenisch}{Mirhosseini et~al.}{2017}]{mirhosseini2017binochs}
\textsc{Mirhosseini, A., M.~Sadrosadati, B.~Soltani, H.~Sarbazi-Azad, and T.~F. Wenisch} (2017): \enquote{BiNoCHS: Bimodal network-on-chip for CPU-GPU heterogeneous systems,} in \emph{Proceedings of the Eleventh IEEE/ACM International Symposium on Networks-on-Chip}, 1--8.

\bibitem[\protect\citeauthoryear{Mittal and Schulz}{Mittal and Schulz}{2013}]{mittal2013general}
\textsc{Mittal, S. and A.~S. Schulz} (2013): \enquote{A general framework for designing approximation schemes for combinatorial optimization problems with many objectives combined into one,} \emph{Operations Research}, 61, 386--397.

\bibitem[\protect\citeauthoryear{Moore}{Moore}{1998}]{moore1998cramming}
\textsc{Moore, G.~E.} (1998): \enquote{Cramming more components onto integrated circuits,} \emph{Proceedings of the IEEE}, 86, 82--85.

\bibitem[\protect\citeauthoryear{Navon and Gopher}{Navon and Gopher}{1979}]{navon1979economy}
\textsc{Navon, D. and D.~Gopher} (1979): \enquote{On the economy of the human-processing system.} \emph{Psychological review}, 86, 214.

\bibitem[\protect\citeauthoryear{Norman and Bobrow}{Norman and Bobrow}{1975}]{norman1975data}
\textsc{Norman, D.~A. and D.~G. Bobrow} (1975): \enquote{On data-limited and resource-limited processes,} \emph{Cognitive psychology}, 7, 44--64.

\bibitem[\protect\citeauthoryear{Palma, Segovia, Kassas, Ribera, and Hall}{Palma et~al.}{2018}]{palma2018self}
\textsc{Palma, M.~A., M.~S. Segovia, B.~Kassas, L.~A. Ribera, and C.~R. Hall} (2018): \enquote{Self-control: Knowledge or perishable resource?} \emph{Journal of Economic Behavior \& Organization}, 145, 80--94.

\bibitem[\protect\citeauthoryear{Pashler}{Pashler}{1994}]{pashler1994dual}
\textsc{Pashler, H.} (1994): \enquote{Dual-task interference in simple tasks: data and theory.} \emph{Psychological bulletin}, 116, 220.

\bibitem[\protect\citeauthoryear{Powell}{Powell}{2008}]{powell2008quantum}
\textsc{Powell, J.~R.} (2008): \enquote{The quantum limit to Moore's law,} \emph{Proceedings of the IEEE}, 96, 1247--1248.

\bibitem[\protect\citeauthoryear{Real, Aggarwal, Huang, and Le}{Real et~al.}{2019}]{real2019regularized}
\textsc{Real, E., A.~Aggarwal, Y.~Huang, and Q.~V. Le} (2019): \enquote{Regularized evolution for image classifier architecture search,} in \emph{Proceedings of the aaai conference on artificial intelligence}, vol.~33, 4780--4789.

\bibitem[\protect\citeauthoryear{Sekeris and Siqueira}{Sekeris and Siqueira}{2024}]{sekeris2024conflict}
\textsc{Sekeris, P.~G. and K.~Siqueira} (2024): \enquote{Conflict and returns to scale in production,} \emph{Journal of Economic Behavior \& Organization}, 227, 106735.

\bibitem[\protect\citeauthoryear{Simon}{Simon}{1955}]{simon1955behavioral}
\textsc{Simon, H.~A.} (1955): \enquote{A behavioral model of rational choice,} \emph{The quarterly journal of economics}, 99--118.

\bibitem[\protect\citeauthoryear{Sims}{Sims}{2003}]{sims2003implications}
\textsc{Sims, C.~A.} (2003): \enquote{Implications of rational inattention,} \emph{Journal of monetary Economics}, 50, 665--690.

\bibitem[\protect\citeauthoryear{Snoek, Larochelle, and Adams}{Snoek et~al.}{2012}]{snoek2012practical}
\textsc{Snoek, J., H.~Larochelle, and R.~P. Adams} (2012): \enquote{Practical bayesian optimization of machine learning algorithms,} \emph{Advances in neural information processing systems}, 25.

\bibitem[\protect\citeauthoryear{Syed, Suriadi, Adams, Bandara, Leemans, Ouyang, Ter~Hofstede, Van De~Weerd, Wynn, and Reijers}{Syed et~al.}{2020}]{syed2020robotic}
\textsc{Syed, R., S.~Suriadi, M.~Adams, W.~Bandara, S.~J. Leemans, C.~Ouyang, A.~H. Ter~Hofstede, I.~Van De~Weerd, M.~T. Wynn, and H.~A. Reijers} (2020): \enquote{Robotic process automation: contemporary themes and challenges,} \emph{Computers in Industry}, 115, 103162.

\bibitem[\protect\citeauthoryear{Thaler}{Thaler}{2016}]{thaler2016behavioral}
\textsc{Thaler, R.~H.} (2016): \enquote{Behavioral economics: Past, present, and future,} \emph{American economic review}, 106, 1577--1600.

\bibitem[\protect\citeauthoryear{Tuk, Zhang, and Sweldens}{Tuk et~al.}{2015}]{tuk2015propagation}
\textsc{Tuk, M.~A., K.~Zhang, and S.~Sweldens} (2015): \enquote{The propagation of self-control: Self-control in one domain simultaneously improves self-control in other domains.} \emph{Journal of Experimental Psychology: General}, 144, 639.

\bibitem[\protect\citeauthoryear{Varian}{Varian}{1992}]{varian1992microeconomic}
\textsc{Varian, H.~R.} (1992): \enquote{Microeconomic analysis,} .

\bibitem[\protect\citeauthoryear{Vaswani}{Vaswani}{2017}]{vaswani2017attention}
\textsc{Vaswani, A.} (2017): \enquote{Attention is all you need,} \emph{Advances in Neural Information Processing Systems}.

\bibitem[\protect\citeauthoryear{Waldrop}{Waldrop}{2016}]{waldrop2016chips}
\textsc{Waldrop, M.~M.} (2016): \enquote{The chips are down for Moore’s law,} \emph{Nature News}, 530, 144.

\bibitem[\protect\citeauthoryear{Westby and Riedl}{Westby and Riedl}{2023}]{westby2023collective}
\textsc{Westby, S. and C.~Riedl} (2023): \enquote{Collective intelligence in human-AI teams: A Bayesian theory of mind approach,} in \emph{Proceedings of the AAAI Conference on Artificial Intelligence}, vol.~37, 6119--6127.

\bibitem[\protect\citeauthoryear{Zamfirescu-Pereira, Wong, Hartmann, and Yang}{Zamfirescu-Pereira et~al.}{2023}]{zamfirescu2023johnny}
\textsc{Zamfirescu-Pereira, J., R.~Y. Wong, B.~Hartmann, and Q.~Yang} (2023): \enquote{Why Johnny can’t prompt: how non-AI experts try (and fail) to design LLM prompts,} in \emph{Proceedings of the 2023 CHI Conference on Human Factors in Computing Systems}, 1--21.\end{thebibliography}
